---
title: "Assignment 9: Visualizing Time Series Data in R"
---

Q1. Collect your own time series data and import into R (e.g. you can use quantmod to collect stock data).

```{r}
# Install and load required packages
options(repos = c(CRAN = "https://cloud.r-project.org/"))

install.packages("quantmod")
library(quantmod)

# Define the stock symbol and the date range
stock_symbol <- "AAPL"
start_date <- as.Date("2022-01-01")
end_date <- Sys.Date()

# Download stock data using getSymbols
getSymbols(stock_symbol, src = "yahoo", from = start_date, to = end_date)

# Extract the adjusted closing prices
stock_data <- Cl(get(stock_symbol))

# Print the first few rows of the data
head(stock_data)


# Plot the chart
chartSeries(AAPL, type = c("auto", "candlesticks", "matchsticks", "bars", "line"), theme = "white")


```

Q2. Examine class of time series object and variables.

```{r}
# Examine the class of the time series object
class(stock_data)

# Examine the structure of the time series object
str(stock_data)
```

Q3. Generate time series charts using tsstudio01.R and dygraphs01.R (in Teams)

```{r}
# Plotting time series data using TSstudio
# install.packages(c("quantmod", "tidyverse","TSstudio"))
# lapply(c("quantmod", "tidyverse","TSstudio"), require, character.only = TRUE)

library(TSstudio)
quantmod::getSymbols("AAPL", src="yahoo")
class(AAPL)
ts_plot(AAPL$AAPL.Adjusted, 
        title = "Apple Stock prices",
        Ytitle = "")
class(AAPL$AAPL.Adjusted) # What class is this object?

# Some sample dataset from TSstudio
ts_seasonal(AAPL, type = "") # month-year matrix data
```

```{r}
library(dygraphs)
par(family="Palatino")
quantmod::getSymbols("AAPL", src="yahoo")

class(AAPL)
m = tail(AAPL, n=30)
m =m[,1:(ncol(m)-2)] # drop last two columns 
names(m)<-c('Open', 'High', 'Low', 'Close') # rename columns for plotting
path <- getwd()
setwd("C:/Users/shiva/Box/My desktop UTD folder backup/UTD/Fall 2023/Data Visualization/Assignments") # place dygraph.css into the same directory
dygraph(m, main = "Apple Stock Prices (Candlestick Chart)")  |>  
  dyCandlestickGroup(c('Open', 'High', 'Low', 'Close')) |> 
  dyCandlestick()  |> 
  dyLegend(show = "always", hideOnMouseOut = T) |> 
  dyCSS("dygraph.css")

```

a\. Analyze time series variables: i. Trend ii. Stationarity iii. pdq

i\. Trend: Trend refers to the general movement of the data over time. The time series plot shows us that the overall trend of the Apple stock data had a fairly stable neutral trend From Januray 2022 until November 2022, which has since been on a general upwards trend.

ii\. Stationarity: To check for stationarity, we plot the time series data and check for no specific trend, constant mean and variance and nonseasonality. Since the Apple stock price data shows a general positive trend, it is nonstationary.

iii\. ARIMA Model Parameters (pdq): AutoCorrelation and Partial AutoCorrelation Function (ACF and PACF): The grid search approach and visual inspection are essential steps in finding appropriate values for p and q. The selected model should be based on both statistical measures and practical considerations. The auto.arima function automatically searches for the best ARIMA model based on AIC (Akaike Information Criterion). The function considers different combinations of p, d, and q and selects the model with the lowest AIC value -- for this case, chosen as, ARIMA (5,2,0) as shown in code below.

```{r}
install.packages("forecast")
library(forecast)

# Grid search for ARIMA
auto_arima_model <- auto.arima(AAPL$AAPL.Adjusted)
summary(auto_arima_model)


```
