---
title: "Assignment 2: Recreating Paul Murrell’s RGraphics basic R programs"
---

1.  Run Paul Murrell\'s RGraphics basic R programs line by line and address the questions if there is one in comments.

    \(i\) Can you change pch?

    Yes, you can change the pch. In this case, **`pch=16`** specifically represents a solid circle. The **`pch`** parameter allows you to customize the appearance of the points in the plot. By changing the value of **`pch`**, you can use different symbols or characters to represent the data points.

    For example:

    -   **`pch=1`** represents a point.

    -   **`pch=2`** represents an open circle.

    -   **`pch=3`** represents a closed triangle.

    -   **`pch=4`** represents an open triangle.

    -   and so on.

    ```{r}
    ## Start plotting from basics 
    # Note the order
    plot(pressure, pch=16)  # Can you change pch? 
    text(150, 600, 
         "Pressure (mm Hg)\nversus\nTemperature (Celsius)")

    ```

\(ii\) Scatterplot: Made after changing the cex from 2 to 3, and the background color of points in x, y2 line from white to blue.

axis(1, at=seq(0, 16, 4)) \# What is the first number standing for?

In the R code **`axis(1, at=seq(0, 16, 4))`**, the first number, which is **`1`**, indicates that the axis being modified is the x-axis. The **`axis()`** function in R is used to add axis labels or tick marks to a plot.

The syntax for **`axis()`** is **`axis(side, at, labels, ...)`** where:

-   **`side`** specifies which axis to draw (**`1`** for x-axis, **`2`** for y-axis, **`3`** for top x-axis, and **`4`** for right y-axis).

-   **`at`** specifies the positions at which tick marks should be drawn.

-   **`labels`** can be used to provide labels for the tick marks.

So, in the context of **`axis(1, at=seq(0, 16, 4))`**, it means that tick marks are being added to the x-axis at positions specified by **`seq(0, 16, 4)`**. The **`seq(0, 16, 4)`** generates a sequence of numbers from 0 to 16 with a step size of 4, so the tick marks will be placed at 0, 4, 8, 12, and 16 on the x-axis.

```{r}
# Scatterplot
# Note the incremental additions

x <- c(0.5, 2, 4, 8, 12, 16)
y1 <- c(1, 1.3, 1.9, 3.4, 3.9, 4.8)
y2 <- c(4, .8, .5, .45, .4, .3)

# Setting label orientation, margins c(bottom, left, top, right) & text size
par(las=1, mar=c(4, 4, 2, 4), cex=.7) 
plot.new()
plot.window(range(x), c(0, 6))
lines(x, y1)
lines(x, y2)
points(x, y1, pch=16, cex=2) # Try different cex value? Changed cex from 3 to 2  
points(x, y2, pch=21, bg="blue", cex=3)  # Different background color: changing it from white to blue
par(col="gray50", fg="gray50", col.axis="gray50")
axis(1, at=seq(0, 16, 4)) # What is the first number standing for?
axis(2, at=seq(0, 6, 2))
axis(4, at=seq(0, 6, 2))
box(bty="u")
mtext("Travel Time (s)", side=1, line=2, cex=0.8)
mtext("Responses per Travel", side=2, line=2, las=0, cex=0.8)
mtext("Responses per Second", side=4, line=2, las=0, cex=0.8)
text(4, 5, "Bird 131")
par(mar=c(5.1, 4.1, 4.1, 2.1), col="black", fg="black", col.axis="black")
```

\(iii\) Histogram:

```{r}
# Histogram
# Random data
Y <- rnorm(50)
# Make sure no Y exceed [-3.5, 3.5]
Y[Y < -3.5 | Y > 3.5] <- NA # Selection/set range
x <- seq(-3.5, 3.5, .1)
dn <- dnorm(x)
par(mar=c(4.5, 4.1, 3.1, 0))
hist(Y, breaks=seq(-3.5, 3.5), ylim=c(0, 0.5), 
     col="gray80", freq=FALSE)
lines(x, dnorm(x), lwd=2)
par(mar=c(5.1, 4.1, 4.1, 2.1))

```

\(iv\) Barplot:

This R code below from Murrell generates a side-by-side boxplot comparing the tooth length (**`len`**) for different doses of either ascorbic acid ("VC") or orange juice ("OJ") in the **`ToothGrowth`** dataset.

```{r}
# Boxplot
par(mar=c(3, 4.1, 2, 0))
boxplot(len ~ dose, data = ToothGrowth,
        boxwex = 0.25, at = 1:3 - 0.2,
        subset= supp == "VC", col="white",
        xlab="",
        ylab="tooth length", ylim=c(0,35))
mtext("Vitamin C dose (mg)", side=1, line=2.5, cex=0.8)
boxplot(len ~ dose, data = ToothGrowth, add = TRUE,
        boxwex = 0.25, at = 1:3 + 0.2,
        
        subset= supp == "OJ")
legend(1.5, 9, c("Ascorbic acid", "Orange juice"), 
       fill = c("white", "gray"), 
       bty="n")
par(mar=c(5.1, 4.1, 4.1, 2.1))
```

\(v\) Perspective plot:

This R code generates a 3D perspective plot of a mathematical function using the persp() function. The function is defined by f(x, y) over a specified range of x and y values. The plot is configured with specific viewing angles and margins.

```{r}
# Perspective plot
x <- seq(-10, 10, length= 30)
y <- x
f <- function(x,y) { r <- sqrt(x^2+y^2); 10 * sin(r)/r }
z <- outer(x, y, f)
z[is.na(z)] <- 1
# 0.5 to include z axis label
par(mar=c(0, 0.5, 0, 0), lwd=0.5)
persp(x, y, z, theta = 30, phi = 30, 
      expand = 0.5)
par(mar=c(5.1, 4.1, 4.1, 2.1), lwd=1)
```

\(vi\) Piechart:

This R code is creating a pie chart using the pie() function to visualize sales distribution for different pie flavors. It sets the margins, character expansion factor. The cex=0.5 parameter reduces the size of the chart elements for better readability. Then it creates a vector specifying the sales distribution for each pie flavor, and then creates a pie chart using pie() fucntion and specifying the color of each pie slice as different shades of gray!

```{r}
# Piechart
par(mar=c(0, 2, 1, 2), xpd=FALSE, cex=0.5)
pie.sales <- c(0.12, 0.3, 0.26, 0.16, 0.04, 0.12)
names(pie.sales) <- c("Blueberry", "Cherry",
                      "Apple", "Boston Cream", "Other", "Vanilla")
pie(pie.sales, col = gray(seq(0.3,1.0,length=6))) 
```

2.  Happy Planet Data:

    ```{r}
    # Install and load required packages
options(repos = c(CRAN = "https://cloud.r-project.org/"))
    install.packages("readxl")
    ```
# Read data from Excel file
happy_data <- read_excel("happy_planet_data.xlsx", sheet = "Sheet1")

# Scatterplot
par(mfrow=c(2, 3))  
plot(happy_data$GDP, happy_data$Life.Expectancy, main="Scatterplot", xlab="GDP", ylab="Life Expectancy")

# Histogram
hist(happy_data$Life.Expectancy, main="Histogram", xlab="Life Expectancy", col="lightblue")

# Barplot
barplot(happy_data$Wellbeing, names.arg=happy_data$Country, main="Barplot", xlab="Country", ylab="Wellbeing", las=2, col="skyblue")

# Perspective Plot
persp(happy_data$GDP, happy_data$Life.Expectancy, happy_data$Wellbeing, main="Perspective Plot", xlab="GDP", ylab="Life Expectancy", zlab="Wellbeing", theta=30, phi=20)

# Pie Chart
pie(happy_data$Happy.Planet.Index, labels=happy_data$Country, main="Pie Chart", col=rainbow(length(happy_data$Country)))
```
```{r}

